\subsection{Problem Definition and Objectives}

In this project, we aim to design and implement a distributed authentication and authorization system for medical queries, called Medchain. We also intend to make this system specifically compatible with the ecosystem of \href{https://medco.epfl.ch/}{MedCo}. MedCo is an existing technology that enables privacy-preserving analysis and sharing of medical data. MedCo already has authentication and authorization mechanisms in place, however, they have some drawbacks which Medchain tries to address. First, in the current implementation of MedCo, both authentication and authorization are centralized. This introduces a single point of failure in the whole ecosystem and has to be avoided. Second, hospitals and medical institutions that participate in MedCo cluster have their own identity providers and run access management protocols individually as these processes can not be done in a federated manner. Our proposed software system, Medchain, offers solutions for such shortcomings. It also puts forward other sought-after properties, such as auditability by recording all the queries submitted to it in a blockchain, which is explained in more details later in this report. 


\subsection{Report Outline}
In this report, we first provide some background information on the underlying technologies used in Medchain deployment in Section \ref{section2}. Next, in Section \ref{section3}, design and architecture of Medchain is presented and discussed. In Section \ref{section4}, we give a full descriptions of our solution, its software design and building blocks. Later, in Section \ref{section5}, we discuss the performance of our software solution and its limitations. In Section \ref{section6}, we discuss the future opportunities to improve the solution and, finally, we finish with conclusions in Section \ref{section7}.